---
title: "Task 3 - Seabirds"
output: html_notebook
---

```{r}
library(tidyverse)
library(here)

seabirds <- read_csv(here("clean_data/seabirds_cleaned.csv"))
```

## 1. Which bird had the most individual sightings?
```{r}
seabirds %>%
  group_by(common_name, scientific_name, abbreviation) %>%
  summarise(total_sightings = n_distinct(id)) %>%
  ungroup() %>%
  slice_max(total_sightings, n = 1, with_ties = TRUE)
```
Wandering albatross sensu lato PL2	Diomedea antipodensis / exulans PL2 DIOANTEXU PL2	3099

## 2. Which bird had the highest total count?
```{r}
seabirds %>% 
  group_by(common_name, scientific_name, abbreviation) %>%
  summarise(total_count = sum(count)) %>%
  ungroup() %>%
  slice_max(total_count, n = 1, with_ties = TRUE)
```
Short-tailed shearwater	Puffinus tenuirostris	PUFTEN	982553


## 3. Which bird had the highest total count above a latitude of -30?
```{r}
seabirds %>%
  group_by(common_name, scientific_name, abbreviation) %>%
  filter(lat > -30) %>%
  summarise(total_count = sum(count)) %>%
  ungroup() %>%
  slice_max(total_count, n = 1, with_ties = TRUE)
```
Wedge-tailed shearwater	Puffinus pacificus	PUFPAC	855


## 4. How many different types of birds were only ever seen in groups of 1?
```{r}
seabirds %>%
  group_by(common_name) %>%
  summarise(max_group_seen = sum(count) / n_distinct(record_id)) %>%
  filter(max_group_seen == 1) %>%
  ungroup() %>%
  summarise(groups_one_only = sum(max_group_seen))
```
86 in groups of one only


## 5. How many penguins were seen? (Hint: there are many types of penguin)
```{r}
seabirds %>%
  filter(str_detect(common_name, regex("penguin", ignore_case = TRUE))) %>%
  summarise(number_of_species = n_distinct(common_name),
            total_number_penguins = sum(count))
```
158
