---
title: "Task 2 - Cake Ingredients"
output: html_notebook
---
```{r}
library(tidyverse)
library(here)
```

```{r}
# Read in cleaned data
cake_ingredients <- read_csv(here("clean_data/cake_ingredients_cleaned.csv"))
```


## 1. Which cake has the most cocoa in it?
```{r}
cake_ingredients %>%
  filter(ingredient == "Cocoa") %>%
  slice_max(quantity, n = 1, with_ties = TRUE)
```
One Bowl Chocolate cake - 10 tablespoons


## 2. For sponge cake, how many cups of ingredients are used in total?
```{r}
cake_ingredients %>%
  group_by(cake) %>%
  filter(cake == "Sponge", measure == "cup") %>%
  summarise(total_quantity_cups = sum(quantity))

```
Sponge cake has 3.5 cups of ingredients

## 3. How many ingredients are measured in teaspoons?
```{r}
cake_ingredients %>%
  group_by(measure) %>%
  filter(measure == "teaspoon") %>%
  summarise(ingredient_count = n_distinct(ingredient))
```
There are 8 ingredients measured in teaspoons.


## 4. Which cake has the most unique ingredients?
```{r}
cake_ingredients %>%
  group_by(cake) %>%
  summarise(number_of_ingredients = n_distinct(ingredient)) %>%
  slice_max(number_of_ingredients, n = 1, with_ties = TRUE)
```
Both Babas au Rhum and One Bowl Chocolate cake are tied and have 11 ingredients


## 5. Which ingredients are used only once?
```{r}
cake_ingredients %>%
  count(ingredient) %>%
  filter(n == 1)
```

Bananas, Cream of tartar, crushed ice, dried currants, egg white, nutmeg, nuts and zwiebach are only used once

