---
title: "Task 4 - Boing Boing Candy Data"
output: html_notebook
---
```{r}
library(tidyverse)
library(here)
```

```{r}
candy_data <- read_csv(here("clean_data/candy_data_cleaned.csv"), col_types = "ciiccccc")
```


## 1. What is the total number of candy ratings given across the three years. (number of candy ratings, not number of raters. Don’t count missing values)
```{r}
candy_data %>%
  filter(!is.na(rating)) %>%
  summarise(total_num_ratings = nrow(.))
```


## 2. What was the average age of people who are going out trick or treating and the average age of people 3. not going trick or treating
```{r}
candy_data %>%
  filter(!is.na(age),
         !is.na(trick_or_treat),
         age > 1 & age < 100) %>%
  group_by(trick_or_treat) %>%
  summarise(average_age = round(mean(age), digits = 1))
```


## 3. For each of joy, despair and meh, which candy bar revived the most of these ratings?
```{r}
candy_data %>%
  group_by(rating, candy_name) %>%
  summarise(number_ratings = n()) %>%
  slice_max(number_ratings, with_ties = TRUE)
  
```

## 4. How many people rated Starburst as despair?
```{r}
candy_data %>%
  filter(candy_name == "Starburst",
         rating == "DESPAIR") %>%
  summarise(starburst_despair_count = n())
```
## For the next three questions, count despair as -1, joy as +1 and meh as 0.

```{r}
candy_rating_int <- candy_data %>%
  mutate(rating = as.integer(case_when(
    rating == "JOY" ~ 1,
    rating == "MEH" ~ 0,
    rating == "DESPAIR" ~ -1)
    )
  )
```


## 5. What was the most popular candy bar by this rating system for each gender in the dataset?
```{r}
candy_rating_int %>%
  filter(!is.na(gender)) %>%
  group_by(gender, candy_name) %>%
  summarise(total_rating = sum(rating)) %>%
  slice_max(total_rating, n = 1, with_ties = TRUE)
```

## 6. What was the most popular candy bar in each year?
```{r}
candy_rating_int %>%
  group_by(year, candy_name) %>%
  summarise(total_rating = sum(rating)) %>%
  slice_max(total_rating, n = 1, with_ties = TRUE)
```


## 7. What was the most popular candy bar by this rating for people in US, Canada, UK and all other countries?
```{r}
candy_rating_int %>%
  drop_na(country) %>%
  mutate(country = case_when(
    str_detect(country, "United Kingdom|United States|Canada") ~ country,
    TRUE ~ "Other")
  ) %>%
  group_by(country, candy_name) %>%
  summarise(total_rating = sum(rating)) %>%
  slice_max(total_rating, n = 1, with_ties = TRUE)
```

```{r}
candy_data %>%
  group_by(country) %>%
  summarise(total_raters = n_distinct(id))

```

