---
title: "Task 1 - Decathlon Data"
output: html_notebook
---


```{r}
library(tidyverse)
library(here)
```

```{r}
decathlon_data <- read_csv(here("clean_data/decathlon_cleaned.csv"))
```


## 1. Who had the longest long jump seen in the data?

```{r}
decathlon_data %>%
  select(id, name, long_jump) %>%
  slice_max(long_jump, with_ties = TRUE)
```
Clay had the longest long jump in this dataset.


## 2. What was the average 100m time in each competition?

```{r}
decathlon_data %>%
  group_by(competition) %>%
  summarise(avg_100m_time = mean(x100m))
  
```
Average 100m time for Decastar was 11.2s and OlympicG 10.9s


## 3. Who had the highest total points across both competitions?

```{r}
decathlon_data %>%
  group_by(name) %>%
  summarise(total_points = sum(points)) %>%
  slice_max(total_points, with_ties = TRUE)
```
Sebrle with 17110 total points


## 4. What was the shot-put scores for the top three competitors in each competition?

```{r}
decathlon_data %>%
  group_by(competition) %>%
  slice_min(rank, n = 3) %>%
  select(name, competition, rank, shot_put)
```

## 5. What was the average points for competitors who ran the 400m in less than 50 seconds vs. those than ran 400m in more than 50 seconds?

```{r}
decathlon_data %>%
  mutate(time_400m = if_else(x400m < 50, "Less than 50s", "More than 50s")) %>%
  group_by(time_400m) %>%
  summarise(avg_points = mean(points))
```

