---
title: "Task 1 - Decathlon Data"
author: "Lee McDonald - DE9"
output: 
  html_document:
    toc: true
    toc_float: true
    number_sections: true
    df_print: paged
  pdf_document: default
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = F)
```

# Introduction

For this task we've been given a dataset of decathlon data, containing athletes' scores on different events and their overall points/rank in the competition.

At first glance, the data doesn't look bad as it is complete with no missing values, however there are some things we can do to make it more "tidy" and suitable for our analysis.

<br>

# Assumptions

To answer the questions we have been asked, I will be removing event data not relevant to the events **100m**, **400m**, **Long jump** and **Shot-put**.

I have also assumed as this is a small dataset, that the recurring names between the Decastar" and "OlympicG" competitions are the same atheletes.

<br>

# Data Cleaning

I used the packages **tidyverse**, **janitor** and **here** to clean this dataset.

The first step I took was to remove the row names and create a new column `name` using `rownames_to_column()`. I then created a unique `athlete_id` for each athlete, by first changing all names to the same case, grouping by `name`, then using `cur_group_id()` which generates an id for each value in the current group.

```{r eval=FALSE}
decathlon_cleaned <- decathlon_data %>%
  # Change row names to column called "name"
  rownames_to_column(var = "name") %>%
  # Create unique id for each athlete
  mutate(name = str_to_title(name)) %>%
  group_by(name) %>%
  mutate(athlete_id = cur_group_id()) %>%
  ungroup() %>%
```

Next, I used `select()` to rearrange the columns and drop any event columns that we did not require for analysis. I then used `pivot_longer()` to change the data to long format, with each observation being an athlete's `score` in an `event`.

```{r eval=FALSE}
  # rearrange columns and drop columns not needed for analysis
  select(athlete_id, name, Competition, Points, Rank, Long.jump, "100m", Shot.put, "400m") %>%
  # Convert into long format
  pivot_longer(Long.jump:"400m", names_to = "event", values_to = "score") %>%
```

Now that the data is in long format, I created a unique `id` for each observation using `row_number()`, and then used `select()` to bring this to the first column.

```{r eval=FALSE}
  # Create id for each observation,
  mutate(id = row_number()) %>%
  # Move id column to front
  select(id, everything()) %>%
```

Finally, I used `clean_names()` to tidy up the remaining column names, and used `str_replace()` to remove the "." character that was in some of the `event` names and replace it with a space.

```{r eval=FALSE}
  # clean column names
  clean_names() %>%
  # replace "." in event names with a space
  mutate(event = str_replace(event, "\\.", " "))
```

<br>

# Analysis

```{r}
library(tidyverse)
library(here)

# Read in clean data
decathlon_data <- read_csv(here("clean_data/decathlon_cleaned.csv"))

# Preview data
decathlon_data
```

<br>

## Question 1 
**Who had the longest long jump seen in the data?**

First I selected the columns we needed, then filtered for "Long jump" in the `event` column, and used `slice_max()` to pick the top result.

```{r}
decathlon_data %>%
  select(name, event, score) %>%
  filter(event == "Long jump") %>%
  slice_max(score, with_ties = TRUE)
```
<br>

## Question 2 
**What was the average 100m time in each competition?**

First `group_by()` for `competition`, then `filter()` for only 100m scores, and `summarise()` to find the `mean()` `score` for each competition.
```{r}
decathlon_data %>%
  group_by(competition) %>%
  filter(event == "100m") %>%
  summarise(avg_100m_time = mean(score))
  
```
<br>


## Question 3 
**Who had the highest total points across both competitions?**

Here I first selected the columns `name`, `points` and `competition` and used distinct() to return only one row of an athlete's `points` per `competition`. I then grouped by `name`, and used `summarise()` to find the `sum()` of each athlete's `points.` Finally `slice_max()` picked the top scoring athlete.

```{r}
decathlon_data %>%
  select(name, points, competition) %>%
  distinct() %>%
  group_by(name) %>%
  summarise(total_points = sum(points)) %>%
  slice_max(total_points, n = 1, with_ties = TRUE)
```
<br>


## Question 4 
**What was the shot-put scores for the top three competitors in each competition?**

First, I grouped by `competition`, then used `filter()` to return only "Shot put" scores in `event`. 

I then used `slice_min()` to find the top three ranked athletes, and finally `select()` to return `name`, `competition`, `rank` and `score`.
```{r}
decathlon_data %>%
  group_by(competition) %>%
  filter(event == "Shot put") %>%
  slice_min(rank, n = 3) %>%
  select(name, competition, rank, score)
```

<br>

## Question 5 
**What was the average points for competitors who ran the 400m in less than 50 seconds vs. those than ran 400m in more than 50 seconds?**

To find this, I filtered to return only rows for the "400m" `event`, then used `mutate()` and `if_else()` to create a new column `time_400m` with the value "**Less than 50s**" or "**More than 50s**".

I then grouped by the new `time_400m` column, and used `summarise()` to find the `mean()` points of these groups.

```{r}
decathlon_data %>%
  filter(event == "400m") %>%
  mutate(time_400m = if_else(score < 50, "Less than 50s", "More than 50s")) %>%
  group_by(time_400m) %>%
  summarise(avg_points = mean(points))
```

