---
title: "Task 5 - RWA Data"
output: html_notebook
---


```{r}
library(tidyverse)
library(here)
```

```{r}
rwa_data <- read_csv(here("clean_data/rwa_data_cleaned.csv"))
```

## 1. What’s the average RWA score for each gender?
```{r}
rwa_data %>%
  group_by(gender) %>%
  summarise(avg_rwa_score = round(mean(rwa_score), 2))
```


## 2. What’s the average RWA score for left handed people vs. right handed people.
```{r}
rwa_data %>%
  group_by(hand) %>%
  filter(hand %in% c("1. Right", "2. Left")) %>%
  summarise(avg_rwa_score = round(mean(rwa_score), 2))
```


## 3. What’s the average family size for each type of childhood?
```{r}
rwa_data %>%
  filter(!is.na(family_size)) %>%
  group_by(urban) %>%
  summarise(avg_family_size = round(mean(family_size), 2))
```


## 4. What’s the average time to take the test for each education level?
Perhaps put a reasonable limit on time taken for the absolute anomalies that occur here
```{r}
rwa_avg_time_education <- rwa_data %>%
  filter(!is.na(education),
         time_secs < 100000) %>%
  group_by(education) %>%
  summarise(avg_time_taken = mean(time_secs))
```


## 5. Create a plot of results of question 4.
```{r}
ggplot(rwa_avg_time_education, aes(education, avg_time_taken)) +
  geom_point()
```

## 6. What’s the average RWA score for people aged:
## Under 18, 18 to 25, 26 to 40, 41 to 60, Over 60
```{r}
rwa_data %>%
  mutate(age_category = case_when(
    age < 18 ~ "Under 18",
    age <= 25 ~ "18 to 25",
    age <= 40 ~ "26 to 40",
    age <= 60 ~ "41 to 60",
    TRUE ~ "Over 60")
  ) %>%
  group_by(age_category) %>%
  summarise(avg_rwa_score = round(mean(rwa_score), 2))
```

